---
title: "Application of ERGM Count in Network Representation of Football Matches"
author: "Imaduddin Haetami"
date: "28/01/2017"
output: 
  revealjs::revealjs_presentation:
    theme: white
    center: TRUE
---

```{r, echo=FALSE,message=FALSE}
knitr::opts_chunk$set(fig.align="center",echo=F,warning = F,cache=F,eval =T, message = F)
library(ztable)
library(ggplot2)
table_LS <- readRDS(paste(getwd(),"/processed data/table_LS",sep=""))
```

#Specification of ERGM Count

***
- In ERGM Count, an edge can have any value of natural numbers including zero, such that $y_{ij} \in \mathbb{N}_0$. 
- The ERGM Count of $\mathbf{Y}$ has Probability Mass Function (PMF) of:  $$Pr_{\mathbf{\theta};h,\eta,g}(\mathbf{Y}=\mathbf{y})=\frac{h(\mathbf{y})\exp(\eta(\mathbf{\theta})\cdot g(\mathbf{y}))}{k_{h,\eta,g}(\mathbf{\theta})}$$
- Which the denominator is a normalization constant defined as:
$$k_{h,\eta,g}(\mathbf{\theta}) = \sum_{\mathbf{y} \in \mathcal{Y}}h(\mathbf{y})\exp(\eta(\mathbf{\theta})\cdot g(\mathbf{y}))$$

#Reference Measure and Baseline Distribution

***
```{r geom_pois,fig.asp=0.4,fig.cap= "Geometric and Poisson Distribution with Mean = 1.5"}
geom_pois <- readRDS(paste(getwd(),"/plots/geom_pois",sep=""))
plot(geom_pois)
```



#Reference Measure and Baseline Distribution (Geometric)

***
- Using sufficient statistics, __sum__ : $g(\mathbf{y}) = \sum_{(i,j) \in \mathbb{Y}} y_{ij}$ 
- Geometric reference measure: $h(\mathbf{y}) = 1$ 
- The model has PMF:
$$Pr(\mathbf{Y}=\mathbf{y}|\theta) \propto \exp\left(\theta \sum_{(i,j) \in \mathbb{Y}} y_{ij}\right) = \prod_{(i,j) \in \mathbb{Y}} (\exp(\theta))^{y_{ij}}$$
- This can be identified as __edge-wise distribution__ $Y_{ij}\stackrel{\text{iid}}{\sim}\text{geometric}(p =1-\exp(\theta))$ where $\theta < 0$

***
```{r, results='asis'}
print(table_LS$geometric,type="html")
```



#Reference Measure and Baseline Distribution (Poisson)

***
- Using sufficient statistics, __sum__ : $g(\mathbf{y}) = \sum_{(i,j) \in \mathbb{Y}} y_{ij}$
- Poisson reference measure: $h(\mathbf{y}) = \prod_{(i,j) \in \mathbb{Y}}(y_{ij}!)^{-1}$
- The model has PMF:
$$Pr(\mathbf{Y}=\mathbf{y}|\theta) \propto \frac{\exp\left(\theta \sum_{(i,j) \in \mathbb{Y}} y_{ij}\right)}{\prod_{(i,j) \in \mathbb{Y}}y_{ij}!} = \prod_{(i,j) \in \mathbb{Y}} \frac{(\exp(\theta))^{y_{ij}}}{y_{ij}!}$$
- This can be identified as __edge-wise distribution__ $Y_{ij}\stackrel{\text{iid}}{\sim}\text{poisson}(\lambda =\exp(\theta))$ where $\theta \in \mathbb{R}$. 

***
```{r, results='asis'}
print(table_LS$poisson,type="html")
```


#Poisson Reference Model
***
we will focus on \textit{poisson-reference} ERGM Count without complex constraints: $\mathcal{Y} \in \mathcal{N}_0^{\mathcal{Y}}$ and $h(\bm{y}) = \prod_{(i,j) \in \mathbb{Y}}(y_{ij}!)^{-1}$. 

#Exogenous Covariates

***
$$
g_k(\bm{y})= \sum_{(i,j) \in \mathbb{Y}} y_{ij}x_{ijk}
$$


***
- Let the model sufficient statistics only be exogenous covariate and __sum__
- the PMF will reduce to:
$$
\begin{eqnarray}
Pr(\bm{Y}=\bm{y}|\theta) &\propto \frac{\exp\left(\theta_1 \sum_{(i,j) \in \mathbb{Y}} y_{ij} + \theta_k \sum_{(i,j) \in \mathbb{Y}} y_{ij}x_{ijk} \right)}{\prod_{(i,j) \in \mathbb{Y}}y_{ij}!}\\
&= \prod_{(i,j) \in \mathbb{Y}} \frac{(\exp(\theta_1+\theta_k x_{ijk}))^{y_{ij}}}{y_{ij}!}
\end{eqnarray}
$$
- This can be recognised as an __edge-wise distribution__ $Y_{ij}\stackrel{\text{iid}}{\sim}\text{poisson}(\lambda =\exp(\theta_1 + \theta_k x_{ijk}))$


##Implementation of Exogenous Covariates
- __nodeocov__ : the attribute of the actor where the edge originate, $x_{i}$. 
- __nodeicov__ : the attribute of receiving actor, $x_j$. 
- __edgecov__ : which represent the external edge attribute, $x_{ij}$.

##__nodeocov__
```{r, results='asis'}
print(table_LS$nodeocov,type="html")
```
##__nodeicov__
```{r, results='asis'}
print(table_LS$nodeicov,type="html")
```

##__edgecov__
```{r, results='asis'}
print(table_LS$edgecov,type="html")
```

##combination
```{r, results='asis'}
print(table_LS$model_8,type="html")
```


#Zero Modification

***
The idea behind zero modification is to move away from the baseline distribution with new PMF that has more accurate representation of its $Pr(Y_{ij}=0)$

***
__nonzero__: $g(\bm{y}) = \sum_{(i,j) \in \mathbb{Y}} \mathbb{I}(y_{ij} > 0)$

***
- Let the model sufficient statistics only be __nonzero__ and __sum__
- Then the PMF will reduce to:
$$
\begin{eqnarray}
Pr(\bm{Y}=\bm{y}|\theta_1,\theta_2) &\propto \frac{\exp\left(\theta_1 \sum_{(i,j) \in \mathbb{Y}} y_{ij} + \theta_2 \sum_{(i,j) \in \mathbb{Y}} \mathbb{I}(y_{ij} > 0)\right)}{\prod_{(i,j) \in \mathbb{Y}}y_{ij}!}\\
&=\prod_{(i,j) \in \mathbb{Y}} \frac{\exp \left(\theta_1 y_{ij} + \theta_2 \mathbb{I}(y_{ij} > 0)\right)}{y_{ij}!}
\end{eqnarray}
$$
- This can be recognised as independent edge-wise Zero Modified Poisson

***
```{r, fig.asp=0.5, fig.cap = "Zero Modified Poisson"}
zero_modified<- readRDS(paste(getwd(),"/plots/zero_modified",sep=""))
plot(zero_modified)
```

***
```{r, results='asis'}
print(table_LS$nonzero,type="html")
```




#Dispersion Modelling

***
Poisson distribution has only 1 parameter that affect the distribution. This parameter found to affect the mean and the spread of the distribution. For one particular value of mean, there is only one particular spread poisson can attain. In other words, the spread is fixed and not adjustable. 

***
__CMP__: $g(\bm{y}) = \sum_{(i,j) \in \mathbb{Y}} \log(y_{ij}!)$

***
- Let the model sufficient statistics only be __CMP__ and __sum__
- Then the PMF will reduce to:
$$
\begin{eqnarray}
Pr(\bm{Y}=\bm{y}|\theta_1,\theta_2) &\propto \frac{\exp\left(\theta_1 \sum_{(i,j) \in \mathbb{Y}} y_{ij} + \theta_2 \sum_{(i,j) \in \mathbb{Y}} \log(y_{ij}!)\right)}{\prod_{(i,j) \in \mathbb{Y}}y_{ij}!}\\
&\propto \prod_{(i,j) \in \mathbb{Y}} (\exp(\theta_1))^{y_{ij}} (y_{ij}!)^{\theta_2-1}
\end{eqnarray}
$$
- Hence the model reduced to edge-wise _Conway-Maxwell-Poisson_ __CMP__ distribution $Y_{ij}\stackrel{\text{iid}}{\sim}\text{CMP}(\lambda =\exp(\theta_1),\nu = 1-\theta_2)$ where $\theta_2 \le 1$.

***
```{r, fig.asp=0.5, fig.cap = "Conway-Maxwell-Poisson Distribution"}
cmp<- readRDS(paste(getwd(),"/plots/cmp",sep=""))
plot(cmp)
```

***
```{r, results='asis'}
print(table_LS$cmp,type="html")
```